---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
1. Package loading
```{r }
library(ggplot2)
library(stringr)
library(dplyr)
```
2. Download data
```{r }
Origin_dir <- getwd()
if(!file.exists('repdata_data_activity.zip')){
  download.file('https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip', 'repdata_data_activity.zip')
}

unzip('repdata_data_activity.zip')
```
3. Read and transform data
```{r }
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date, "%Y-%m-%d" )
```
## What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day, plot in a histogram chart
```{r }
activityByday <- activity %>% group_by(date) %>% summarise( stepsPerday = sum(steps, na.rm = TRUE))

ggplot(activityByday, aes(x=stepsPerday)) + geom_histogram(binwidth = 2000, color = "white", fill = "grey")+
labs(title="Histogram of steps per day (with na)",x="Steps/Day", y = "Counts") 

```  
  
2. Calculate and report the mean of the total number of steps taken per day
```{r}
mean(activityByday$stepsPerday, na.rm = TRUE)
```  
3.Calculate and report the median of the total number of steps taken per day
```{r}
median(activityByday$stepsPerday, na.rm = TRUE)
```  

## What is the average daily activity pattern?
1. Plot average steps by interval across the days.
```{r}
activityByinterval <- activity %>% group_by(interval) %>% summarise(avgsteps = mean(steps, na.rm = TRUE))
## qplot(interval, avgsteps , data = activityByinterval, geom = "line", main = 'Average steps per interval', ylab = 'Average Steps')
ggplot(activityByinterval, aes(x=interval, y=avgsteps)) + geom_line( color = "orange", size =1.2 )+
  labs(title="Average steps per interval (5 minutes)",x="interval", y = "Steps")
```

2. What is the maxium steps at what interval
```{r}
max <- activityByinterval  %>% filter ( avgsteps == max(activityByinterval$avgsteps))
max
```

## Imputing missing values
1. Calculate and report the total number of missing values in the dataset. (i.e. the total number of rows with NAs)
```{r}
sum(is.na(activity$steps))
```
2. Create a new dataset that is equal to the original dataset but with the missing data filled with mean of steps of the interval.
```{r}
activity1 <-  activity %>% group_by(interval) %>% mutate(stepsAdjusted = ifelse(is.na(steps), mean(steps, na.rm = TRUE), steps))
```
3. Calculate the total number of steps taken per day, plot in a histogram chart.
```{r}
activityByday1 <- activity1 %>% group_by(date) %>% summarise( stepsPerday = sum(stepsAdjusted, na.rm = TRUE))
## hist(activityByday1$stepsPerday, main = 'Histogram of steps per day (with fills)', xlab = 'steps') 
ggplot(activityByday1, aes(x=stepsPerday)) + geom_histogram(binwidth = 2000, color = "white", fill = "lightblue")+
  labs(title="Histogram of steps per day (fill na)",x="Steps/Day", y = "Counts")

```

4. Calculate and report the mean of the total number of steps taken per day after fill-na
```{r}
mean(activityByday1$stepsPerday, na.rm = TRUE)
```
5. Calculate and report the median of the total number of steps taken per day after fill-na
```{r}
median(activityByday1$stepsPerday, na.rm = TRUE)
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
activity1 <- activity1 %>% mutate(weekday = weekdays(date)) %>% mutate( dayType = ifelse( weekday == 'Sunday' | weekday == 'Saturday', 'Weekend', 'weekday'))
activityByinterval1 <- activity1 %>% group_by(interval, dayType) %>% summarise(avgsteps = mean(steps, na.rm = TRUE))

ggplot(activityByinterval1, aes(x=interval, y=avgsteps)) + geom_line( color = "blue", size =0.5 )+
  labs(title = "Average steps per interval (5 minutes) by day type", x = "interval", y = "Steps")+
  facet_wrap(dayType ~. , ncol = 1)
````


